
\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{array}
\geometry{margin=1in}

\title{Deterministic Random Bit Generators: A Comparison of NIST-Approved Techniques}
\author{
    Tomas Bedoya \\
    Daniel Andres Jaimes Cardenas \\
    Eder Leandro Carbonero Baquero \\
    \textit{Universidad de los Andes, Colombia}
}
\date{\today}

\begin{document}

\maketitle
\begin{abstract}
The following paper centers around several methods of random number generation. It aims to comprehend the essential concepts behind RNG, including pseudo and true random generation. Additionally, it explores different approaches to randomness that are used today that follow the NIST SP 800-90 standards, both reliant on entropy sources as well as algorithmic techniques, and compares their randomness with a series of statistical tests to determine which operate better and suggest appropriate uses for the different methods.

\end{abstract}

\section*{General defintions}
\input{sections/general_definitions}

\section{Introduction}
\input{sections/introduction}

\section{Key Concepts}
\input{sections/key_concepts}

\section{Studied Methods}
The following RNG methods are the ones that will be seen in detail for the rest of this report.
\input{sections/methodologies}

\section{Statistical Testing}
\input{sections/testing}

\section{Results}
\input{sections/results}

\section{Conclusions}
\input{sections/conclusions}

\section*{References}  
\bibliographystyle{plain}

\begin{thebibliography}{9}
\bibitem{kelsey2019} Kelsey, J., Brand\~{a}o, L. T. A., Peralta, R., Booth, J. (2019). \textit{A Proposal for a Beacon Design}. NIST.
\bibitem{kelsey2018} Kelsey, J. (2018). \textit{Randomness Beacon Background and Overview}. NIST.
\bibitem{mechalas2018} Mechalas, D. (2018). \textit{Intel Digital Random Number Generator (DRNG) Software Implementation Guide}.
\bibitem{zhang2021} Zhang, Y., Lo, H.-P., Mink, A., Ikuta, T., Honjo, T., Takesue, H., Munro, W. J. (2021). \textit{A Simple Low-Latency Real-Time Certifiable Quantum Random Number Generator}. NIST.
\bibitem{kelsey2021} Kelsey, J. (2021). \textit{NIST Beacon Operations Overview}.
\end{thebibliography}



\end{document}
